Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2017-11-03 aggregation/agg-basic-lca.tex
afternoon work- cites; semantics; editing

Because the objective of the framework is to validate a computation, we begin by providing a mathematical formulation of the LCA problem.  An LCA computation is performed over a collection of linked unit processes, each of which describes a single operation or reference product.  The task of LCA can be described in two steps: making an inventory of environmental interventions; and determining their environmental impacts.  If the processes are assumed to be linear, the problem can be written in terms of matrix algebra.  There are two common formulations: the ``technology matrix'' approach and the ``input-output'' or Leontief approach \citep{Suh_2002, Pauliuk_2015}.  Although the two models are understood to be equivalent, the input-output approach is computationally more efficient in some applications because it lends itself to iterative solution \citep{Peters_2007}.  That formulation may be stated:
\begin{equation}
s = \mathbf{e}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the \emph{direct requirements matrix}, where each column of values reports the necessary inputs (positive values) and generated non-reference outputs (negative values) per unit of a process's reference output; $B$ is the environmental intervention or emission matrix; and $\mathbf{e}$ is a column vector of characterization factors for the environmental emissions with respect to a given impact assessment method.  In this model, $\mathbf{y}$ is the externally specified final demand, which is the product or service whose delivery is to be assessed, and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $E$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.

The term $(I-A)$ is equivalent to the normalized technology matrix, i.e. the difference between the supply and use tables in classical input-output analysis, $(V'-U)$ \citep{Suh_2010}.  The conditions under which this matrix has an inverse, particularly with respect to different strategies for handling co-production, have been discussed extensively \citep{Suh_2010, Majeau_Bettez_2014}.  Eq.~\ref{eqn:leontief} is visualized in Figure~\ref{fig:lca}. 
\input{fig_lca}

>>>>

Because the objective of the framework is to validate a computation, we begin by providing a mathematical formulation of the LCA problem.  An LCA computation is performed over a collection of linked unit processes, each of which describes a single operation or reference product.  The task of LCA can be described in two steps: making an inventory of environmental interventions; and determining their environmental impacts.  If the processes are assumed to be linear, the problem can be written in terms of matrix algebra.  There are two common formulations: the ``technology matrix'' approach and the ``input-output'' or Leontief approach \citep{Heijungs2002, Pauliuk_2015_framework}.  Although the two models are understood to be equivalent, the input-output approach is computationally more efficient in some applications because it lends itself to iterative solution \citep{Peters_JLCA_2007}.  That formulation may be stated:
\begin{equation}
s = \mathbf{e}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the \emph{direct requirements matrix}, where each column of values reports the necessary inputs (positive values) and generated non-reference outputs (negative values) per unit of a process's reference output; $B$ is the environmental intervention or emission matrix; and $\mathbf{e}$ is a column vector of characterization factors for the environmental emissions with respect to a given impact assessment method.  In this model, $\mathbf{y}$ is the externally specified final demand, which is the product or service whose delivery is to be assessed, and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $E$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.

The term $(I-A)$ is equivalent to the normalized technology matrix, i.e. the difference between the supply and use tables in classical input-output analysis, $(V'-U)$ \citep{Suh_JIE_2010}.  The conditions under which this matrix has an inverse, particularly with respect to different strategies for handling co-production, have been discussed extensively \citep{Majeau_Bettez_2014}.  Eq.~\ref{eqn:leontief} is visualized in Figure~\ref{fig:lca}. 
\input{fig_lca}

### 2017-11-03 aggregation/agg-lca-study.tex
afternoon work- cites; semantics; editing

For clarity, I will call the augmented $\tilde{A}$ and $\tilde{B}$, in which the foreground content always appears to the left of the background content; the first column delivers the study's functional unit; and the background does not depend on the foreground; to be an \emph{LCA foreground study}, for which $\tilde{\mathbf{y}}$ is the canonical functional unit.  This formulation can be applied to the vast majority of LCA case studes, and very likely to all case studies prepared using commercial LCA software.  However, it does not apply for so-called ``integrated hybrid'' models, in which all quadrants contain non-zero cells (for instance, \cite{Gibon_2016}).

>>>>

For clarity, I will call the augmented $\tilde{A}$ and $\tilde{B}$, in which the foreground content always appears to the left of the background content; the first column delivers the study's functional unit; and the background does not depend on the foreground; to be an \emph{LCA foreground study}, for which $\tilde{\mathbf{y}}$ is the canonical functional unit.  This formulation can be applied to the vast majority of LCA case studes, and very likely to all case studies prepared using commercial LCA software.  However, it does not apply for so-called ``integrated hybrid'' models, in which all quadrants contain non-zero cells (for instance, \cite{Gibon_2015}).

### 2017-11-03 aggregation/agg-objective.tex
afternoon work- cites; semantics; editing

The task of computing life cycle impact assessment (LCIA) results from a product system model can be described as an ``aggregation,'' in which the data in large, sparse models are combined through linear algebra operations into a small set of numeric scores.  Although several equivalent methods for computing aggregated LCI and/or LCIA results have been described in the literature \citep{Suh_2005, Peters_2007}, the computational representation of and exchange of product system models, and the generation of aggregated results, remains software-specific.

In the context of this paper, an ``LCA Study'' should be understood as any aggregation of data under consideration, from a single-operation unit process to a full ISO-style product LCA.  The practice of reporting the results of an LCA study implies the existence of two parties with different information: the study author has full knowledge of the product system model; and the reader only has knowledge of what the author has disclosed.  Obviously, the level of detail of the disclosure will determine how useful it is to the reader.  In this context, the reader will want to know the answers to various review objectives.  
At the same time the author may have constraints that limit what she can disclose.  The purpose of the framework is to give form to the disclosure so that the requirements of various review objectives, and restrictions associated with various disclosure contraints, can be evaluated clearly.

>>>>

The task of computing life cycle impact assessment (LCIA) results from a product system model can be described as an ``aggregation,'' in which the data in large, sparse models are combined through linear algebra operations into a small set of numeric scores.  Although several equivalent methods for computing aggregated LCI and/or LCIA results have been described in the literature \citep{Suh2005a, Peters_JLCA_2007}, the computational representation of and exchange of product system models, and the generation of aggregated results, remains software-specific.  
In the context of this paper, an ``LCA Study'' should be understood as any collection of inventory data under consideration, from a single-operation unit process to a full ISO-style product LCA.

Reporting the results of an LCA study implies the existence of two parties with different information: the study author has full knowledge of the product system model; and the reader only has knowledge of what the author has disclosed.  Obviously, the level of detail of the disclosure will determine how useful it is to the reader.  In this context, the reader will want to know the answers to various review objectives.  
At the same time the author may have constraints that limit what can be disclosed.  The main goal of the framework is to give form to the disclosure so that the requirements of various review objectives, and restrictions associated with various disclosure contraints, can be evaluated clearly.

### 2017-11-03 aggregation/agg-pagg.tex
afternoon work- cites; semantics; editing

The author may relax the disclosure constraints by reporting the locations of some or all non-zero entries in $A_{d,priv}$ and $B_{f,priv}$ but not disclosing their values. This would enable a reader to understand \emph{that} a certain background process or emission was included in the model, but not to know \emph{how much} was included.

>>>>

The author may relax the disclosure constraints by reporting the locations of some or all non-zero entries in $A_{d,priv}$ and $B_{f,priv}$ but not disclosing their values. This would enable a reader to understand \emph{that} a certain background process or emission was included in the model, but not to know \emph{how much} was included, which may jointly satisfy review objectives and disclosure constraints in some cases.

### 2017-11-03 aggregation/agg-review.tex
Some morning work
afternoon work- cites; semantics; editing

Together, these lists can be considered a functional definition of the study scope that can be used to conduct a qualitative comparison of multiple studies by different authors.  Background database choice, dataset selection, and version information can all be made available in disclosure item~\ref{itm:bg}, thereby streamlining what is presently a highly interpretive and labor-intensive process of extracting this information from written text.  Similarly, the set of elementary flows explicitly modeled in the study foreground (item~\ref{itm:em}) expose whether any study-specific modeling of direct emissions was performed and what flows were included.

@@

Partial aggregation is already used in practice to conceal private data, but current practices used to generate aggregated results from input model configurations do not provide a way for a reviewer to validate the aggregation.  In contrast, if a study containing confidential information is partitioned into public and private partitions as in Eq.~\ref{eqn:private} and Fig.~\ref{fig:private}, then the results of the computation can be validated directly.  To achieve this, the reviewer must be granted access to both the public and private partitions in order to confirm that the partitioning is valid according to Eq.~\ref{eqn:partition}, and that $\mathbf{b}_{priv}$ was computed correctly.  However, and the private data can still be withheld from the disclosure.

A further advantage to this approach is that the reader can evaluate easily what fraction of the overall impact score is accounted for by the public versus the private portions of the model.  A ``disclosure completeness'' metric $\varphi$ can be defined as the fraction of the overall score which is accounted for by the disclosed portion of the PSM:
\begin{equation}
  \varphi =  \frac{\mathbf{e}^T\cdot(\mathbf{b} - \mathbf{b}_{priv})}{\mathbf{e}^T\cdot\mathbf{b}} = 1 - \frac{\mathbf{e}^T\cdot\mathbf{b}_{priv}}{\mathbf{e}^T\cdot\mathbf{b}}
  \label{eqn:metric}
\end{equation}
where $\mathbf{b}$ is the complete life cycle inventory.  The value of this metric can provide an indication of the level of transparency of the disclosure with respect to a given impact category indicator.  The author may relax the disclosure constraints by reporting the locations of some or all non-zero entries in $A_{d,priv}$ and $B_{f,priv}$ but not disclosing their values.

>>>>

Together, these lists can be considered a functional definition of the study scope that can be used to conduct a qualitative comparison of multiple studies by different authors.  Background database choice, dataset selection, and version information is all made available in disclosure item~\ref{itm:bg}, thereby streamlining what is presently an interpretive and study-specific process of extracting this information from written text.  Similarly, the set of elementary flows explicitly modeled in the study foreground (item~\ref{itm:em}) expose whether any study-specific modeling of direct emissions was performed and what flows were included.

@@

Partial aggregation is already used in practice to conceal private data, but current practices used to generate aggregated results from input model configurations do not provide a way for a reviewer to validate the aggregation.  If a study containing confidential information is split into public and private partitions as in Eq.~\ref{eqn:private} and Fig.~\ref{fig:private}, then the results of the computation can be validated directly by the reviewer.  To achieve this, the reviewer must be granted access to both the public and private partitions in order to confirm that the partitioning is valid according to Eq.~\ref{eqn:partition}, and that $\mathbf{b}_{priv}$ was computed correctly.  However, the private information used in the computation can still be withheld from the disclosure.

A further advantage to this approach is that the reader can evaluate easily what fraction of the overall impact score is accounted for by the public versus the private portions of the model.  A ``disclosure completeness'' metric $\varphi$ can be defined as the fraction of the overall score which is accounted for by the disclosed portion of the PSM:
\begin{equation}
  \varphi =  \frac{\mathbf{e}^T\cdot(\mathbf{b} - \mathbf{b}_{priv})}{\mathbf{e}^T\cdot\mathbf{b}} = 1 - \frac{\mathbf{e}^T\cdot\mathbf{b}_{priv}}{\mathbf{e}^T\cdot\mathbf{b}}
  \label{eqn:metric}
\end{equation}
where $\mathbf{b}$ is the complete life cycle inventory.  The value of this metric can provide an indication of the level of transparency of the disclosure with respect to a given impact category indicator.

### 2017-11-03 aggregation/agg-semantic.tex
afternoon work- cites; semantics; editing

In actuality, while it is easy to reproduce a set of sparse matrices, there is considerable potential for ambiguity
in identifying the background datasets and emissions to which those matrix entries correspond.
It is necessary for the entries in disclosure items~\ref{itm:bg} and \ref{itm:em} to be unambiguous and easy to interpret.  Entries in \ref{itm:bg}, which represent rows in $A_d$, must clearly identify the originating database, version, and configuration, along with the precise dataset, that dataset's reference flow, and the dimension (reference quantity or unit) of that flow.  In addition, the sign of the numeric entry in $A_d$ must be consistent with the implementation of the background dataset.  Similar requirements also hold for entries in \ref{itm:em}, which represent rows in $B_f$.  Current research has revealed widespread challenges to finding consistency on the identities of elementary flows \citep{Edelen_2017}.

In order to achieve the transparency and reproducibility promised by this disclosure framework, it must be possible for study authors and readers to agree on the meaning of those references.  This is often accomplished using linked semantic data \citep{Bizer_2009}, in which an object to be interpreted is signified by a link to a resource on the World Wide Web, typically referred to as a Uniform Resource Identifier (URI) or a hyperlink.  The resource at the end of the link functions as a point of agreement: both study author and data user can follow it from anywhere on the Internet to obtain the same information.  The content pointed to by the hyperlink, along with context provided by \emph{other} references to the same resource, gives meaning to the information and allows it to be curated by the study authors and other members of the community \citep{Khan_2011}.   % Semantic data are structured by reference to knowledge models called ontologies, which describe the relationships among different types of entities \citep{Madin2008}.  Some preliminary descriptions of the entity types involved in LCA computation, and their relationships to one another, have recently been developed \citep{Ciroth_Srocka_2014, Janowicz_WOP_2015, Kuczenski_JCP_2016}.

Before the goal of easily reproducible foreground models can be realized, the linked data foundation must be laid.  Background database providers should ensure that their users have access to \emph{stable semantic references} for publicly available datasets that distinguish database, version, configuration, dataset, and product flow to which the reference refers.  These references should take the form of URIs that can be accessed using any browser and can provide a reader with documentary information describing the dataset, the quantitative dimension of the reference flow, and a mechanism for the reader to obtain access to the data.  Similarly, LCIA method providers should ensure that their methods make use of stable semantic references to elementary flows and contexts.

>>>>

In practice, while it is easy to reproduce a set of sparse matrices, there is considerable potential for ambiguity
in identifying the background datasets and emissions to which those matrix entries correspond.
In order to achieve the transparency and reproducibility promised by this disclosure framework, it must be possible for study authors and readers to agree on the meaning of those references.  In particular, the database configuration, dataset version, reference flow and dimension (quantitative unit of measure) must be unambiguous and easy to interpret.

Agreement about the meaning of concepts is often accomplished using linked semantic data \citep{Bizer_2009}, in which an object to be interpreted is signified by a link to a resource on the World Wide Web, typically referred to as a Uniform Resource Identifier (URI) or a hyperlink.  The resource at the end of the link functions as a point of agreement: both study author and data user can follow it from anywhere on the Internet to obtain the same information.  The content pointed to by the hyperlink, along with context provided by \emph{other} references to the same content, gives meaning to the information and allows it to be curated by the study authors and other members of the community \citep{Khan_2011}.   Semantic data are structured by reference to knowledge models called ontologies, which describe the relationships among different types of entities \citep{Madin2008}.  Some preliminary descriptions of the entity types involved in LCA computation, and their relationships to one another, have recently been developed \citep{Ciroth_Srocka_2014, Janowicz_WOP_2015, Kuczenski_JCP_2016}.

Before the goal of easily reproducible foreground models can be realized, the linked data foundation must be laid.  Background database providers should ensure that their users have access to \emph{stable semantic references} for publicly available datasets that denote a particular activity in a particular database version and configuration.  These references should take the form of URIs that can be accessed using any Internet browser and can provide both author and reader with documentary information describing the referred dataset, the available reference flows and their dimensions, and a mechanism for the reader to obtain access to the exchange data or LCI/LCIA computations for the purposes of model validation.  A dependency reference in disclosure item~\ref{itm:bg} should include both an activity and a reference flow.

Similarly, LCIA method providers should ensure that their methods make use of stable semantic references to elementary flows and contexts.  Current research has revealed widespread challenges to finding consistency on the identities of elementary flows despite the community's longstanding awareness of the problem, and several mutually inconsistent reference flow sets exist \citep{Edelen_2017}.  The conventional understanding of a ``flow'' as comprising a substance and a context together (e.g. methane, emissions to air) causes a combinatorial increase in the number of flows in a database, and also multiplies the potential points of disagreement across sources.  I reiterate Edelen et al.'s (ibid.) suggestion to explicitly regard substances (which can easily include non-material ``flowables'' such as land occupation and transformation) and contexts as independent semantic classes.  In this case, an elementary flow entry in the disclosure item~\ref{itm:em} would require both a flowable and a context.

\endinput

It is necessary for the entries in disclosure items~\ref{itm:bg} and \ref{itm:em} to be unambiguous and easy to interpret.  Entries in \ref{itm:bg}, which represent rows in $A_d$, must clearly identify the originating database, version, and configuration, along with the precise dataset, that dataset's reference flow, and the dimension (reference quantity or unit) of that flow.  In addition, the sign of the numeric entry in $A_d$ must be consistent with the implementation of the background dataset.  Similar requirements also hold for entries in \ref{itm:em}, which represent rows in $B_f$.
